Bootstrapping Multi-atlas Segmentation Using Multiple Automatically Generated Templates for the Segmentation of the Whole Hippocampus and Subfields

### 2013-12-28 paper.Rnw
FEP Bland-altman and volume comparison charts

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{algorithm, color, algpseudocode, amsmath, lrl, geometry, ctable}
\usepackage{titlesec, siunitx, graphicx}
\usepackage[round,authoryear]{natbib}
\usepackage[section]{placeins}   % keep floats in their place.
\usepackage{authblk}             % Listing Author affiliations
\usepackage{rotating}            % rotating figures
\usepackage[hypcap]{caption}
\usepackage{lineno}

@@

ggplot(data=fep.volumes, aes(x=manual, y=volume, colour=method)) + 
  geom_smooth(method="lm", formula=y~x) + 
  geom_point() + 
  geom_text(aes(x=3350,y=y,label=eqn),parse=T,data=fep.lm,
      size=3,hjust=0) +
  xlab(expression("Mean bilateral manual hippocampus volume " (mm^3))) + 
  ylab(expression("Mean bilateral computed hippocampus volume "  (mm^3))) + 
  scale_size_continuous(guide="none") 
@
  \caption{Mean bilateral hippocampus volume as measured by \mb{}  vs. manually
  segmented volumes from the First Episode Patients with Schizophrenia dataset.
  A linear fit line is shown, with shaded region showing standard error. 
  \label{fig:SZ-volumes}}
\end{figure}

\begin{figure}
<<fig:FEP-Bland-Altman, cache=TRUE, dependson='ADNI-volumes-prep', fig.width=7, fig.height=7>>=
fep.volumes$diff = fep.volumes$manual - fep.volumes$volume
fep.volumes$mean = (fep.volumes$manual + fep.volumes$volume)/2
limits = ddply(fep.volumes, c("method"), function (df) { 
  data.frame(
    y = mean(df$diff,na.rm=T) + c(-1.96,0,2) * sd(df$diff,na.rm=T) 
)})

ggplot(fep.volumes, aes(x= mean, y = diff)) + 
  facet_wrap( ~ method, nrow=2, scales="fixed") + 
  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
  geom_point(size=1) +  
  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
             data=limits, size=0.3, alpha=0.5) + 
  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", data = limits,
            hjust=1.1,vjust=-0.8, size=2.5) + 
  xlab(expression('Mean of SNT and automated hippocampal volume ' (mm^3))) + 
  ylab(expression('Difference in manual and automated hippocampal volume ' (mm^3))) + 
  scale_y_continuous(breaks=seq(-10000,4000, 1000)) + 
  scale_x_continuous(breaks=seq(0,10000, 1000)) + 
  theme(legend.direction = "horizontal", legend.position = "bottom", 
        axis.text.y = element_text(angle = 90, hjust=0.5), 
        axis.text = element_text(size = 8))
@
  \caption{Bland-Altman plots comparing subject mean hippocampal volume as
measured in the FEP dataset by \snt{} segmentation and each of the
four automated methods investigated (\freesurfer{}, \fsl{}, \maper{}, \mb{}). The overall
mean difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by
dashed horizontal lines. Linear fit lines are shown for each diagnosis group.
Note, points below the mean difference indicate overestimation of the volume
with respect to the \snt{} volume, and vice versa. }
  \label{fig:FEP-Bland-Altman}
\end{figure}

<<FEP-t-test,cache=TRUE,dependson="FEP-volumes">>=
  fep.cor.test = cor.test(fep.volumes$manual,fep.volumes$MAGeT_Winterburn)
@
\mb{} produces hippocampal volumes that are highly correlated with manual
segmentation volumes (Pearson $r = \Sexpr{fep.cor.test$estimate}$, $t =
\Sexpr{fep.cor.test$statistic}$, $p < \Sexpr{max(0.001, fep.cor.test$p.value)}$;
Figure \ref{fig:SZ-volumes}).

>>>>

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{algorithm, color, algpseudocode, url, amsmath, geometry, ctable}
\usepackage{titlesec, siunitx, graphicx}
\usepackage[round,authoryear]{natbib}
\usepackage[section]{placeins}   % keep floats in their place.
\usepackage{authblk}             % Listing Author affiliations
\usepackage{rotating}            % rotating figures
\usepackage[hypcap]{caption}
\usepackage{lineno}

@@

ggplot(data=fep.volumes, aes(x=manual, y=volume, colour=method)) + 
  geom_smooth(method="lm", formula=y~x) + 
  geom_point() + 
  geom_text(aes(x=3350,y=y,label=eqn),parse=T,data=fep.lm,
      size=3,hjust=0) +
  xlab(expression("Mean manual hippocampus volume " (mm^3))) + 
  ylab(expression("Mean computed hippocampus volume "  (mm^3))) + 
  theme(legend.direction = "horizontal", legend.position = "bottom")
@
  \caption{Mean bilateral hippocampus volume as measured by \mb{}, \fsl{}, and
\freesurfer{} vs. manually segmented volumes from the First Episode Patients with
Schizophrenia dataset. \mb{} is run two ways: with an atlas library consistening
of the five Winterburn atlases, or with five manually segmented FEP subjects as
atlases (not included as subjects). Linear fit lines are shown, with shaded
region showing standard error. 
  \label{fig:SZ-volumes}}
\end{figure}

\begin{figure}
<<fig:FEP-Bland-Altman, cache=TRUE, dependson='fig:FEP-volumes', fig.width=7, fig.height=7>>=
fep.volumes$diff = fep.volumes$manual - fep.volumes$volume
fep.volumes$mean = (fep.volumes$manual + fep.volumes$volume)/2
limits = ddply(fep.volumes, c("method"), function (df) { 
  data.frame(
    y = mean(df$diff,na.rm=T) + c(-1.96,0,2) * sd(df$diff,na.rm=T) 
)})

ggplot(fep.volumes, aes(x= mean, y = diff)) + 
  facet_wrap( ~ method, nrow=2, scales="fixed") + 
  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
  geom_point(size=1) +  
  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
             data=limits, size=0.3, alpha=0.5) + 
  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", data = limits,
            hjust=1.1,vjust=-0.8, size=2.5) + 
  xlab(expression('Mean manual and computed volume ' (mm^3))) + 
  ylab(expression('Difference in manual and computed volume ' (mm^3))) + 
  scale_y_continuous(breaks=seq(-10000,4000, 1000)) + 
  scale_x_continuous(breaks=seq(0,10000, 1000)) + 
  theme(legend.direction = "horizontal", legend.position = "bottom", 
        axis.text.y = element_text(angle = 90, hjust=0.5), 
        axis.text = element_text(size = 8))
@
  \caption{Bland-Altman plots comparing subject mean hippocampal volume as
measured in the FEP dataset by manual segmentation and each of the
four automated methods investigated (\freesurfer{}, \fsl{}, \mb{} with
Winterburn atlases, \mb{} with FEP manual atlases). The overall
mean difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by
dashed horizontal lines. Linear fit lines are shown for each diagnosis group.
Note, points below the mean difference indicate overestimation of the volume
with respect to the manual volume, and vice versa. }
  \label{fig:FEP-Bland-Altman}
\end{figure}

% <<FEP-t-test,cache=TRUE,dependson="FEP-volumes">>=
%   fep.cor.test = cor.test(fep.volumes$manual,fep.volumes$MAGeT_Winterburn)
% @
% \mb{} produces hippocampal volumes that are highly correlated with manual
% segmentation volumes (Pearson $r = \Sexpr{fep.cor.test$estimate}$, $t =
% \Sexpr{fep.cor.test$statistic}$, $p < \Sexpr{max(0.001, fep.cor.test$p.value)}$;
% Figure \ref{fig:SZ-volumes}). 
%
